A Data-Driven Framework for Neural Field Modelling

### 2010-11-15 ltx/first_order_framework/LaTeX_Manuscript.tex
added alternative table

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again applying Shannon's sampling theorem, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

@@

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} In this section, we provide a demonstration of the estimation procedure using the new methods that are described above. The full neural field model was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field can be seen in Figure~\ref{fig:Figure3}(a). Figure~\ref{fig:Figure3}(b) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(c) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis}
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:Figure4}(a) shows the spatial frequency of the neural field and Figure~\ref{fig:Figure4}(b) shows the spatial frequency of the reconstruct neural field from the basis function decomposition. The cut-off frequency of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing.

@@

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(c). The band-limiting effect of the basis function decomposition can be seen when comparing Figure~\ref{fig:Figure4}(a) to Figure~\ref{fig:Figure4}(b).

@@

\subsubsection{Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure the parameters converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming the parameters converged, with the changes in the parameter errors falling to less than $10^{-4}$ after six iterations.

@@

\tiny{\begin{tabular}{lcccccccccccc}
Experiment&\multicolumn{4}{c}{I}&\multicolumn{4}{c}{II}&\multicolumn{4}{c}{III}\\
 Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$\\
\hline\hline
Actual Value&$90$&$-80$&$5$&$0.9$&$100$&$-80$&$5$&$0.9$&$110$&$-80$&$5$&$0.9$\\
Mean Estimate&$103.86$&$-89.05$&$5.23$&$0.923$&$101.75$&$-81.00$&$4.76$&$0.924$&$98.90 $&$-73.13$&$4.37$&$0.925$\\
Std. Dev.&$23.14$&$16.57$&$0.74$&$0.003$&$21.30$&$14.82$&$0.65$&$0.003$&$16.48$& $11.72$&$0.54$&$0.003$\\
Bias (\%)& $15.4$&$11.31$&$4.6$&$2.56$&$1.75$&$1.25$&$4.8$&$2.67$&$10.1$&$ 8.58$&$12.6$&$2.78$  \\
\end{tabular}\label{tab:MexicanHatKernelResults}}

>>>>

Similar considerations need to be made regarding the representation of the dynamic field, $v_t(\mathbf{r})$, using the basis function decomposition. Again, applying Shannon's sampling theorem, the minimum distance between basis functions must satisfy 
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{cy}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

@@

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} In this section, we provide demonstrations of the estimation procedure using the new methods that are described above. The full neural field model was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}, unless otherwise explicitly stated for particular experiments. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field can be seen in Figure~\ref{fig:Figure3}(a). Figure~\ref{fig:Figure3}(b) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(c) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis}
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:Figure4}(a) shows the spatial frequency of the neural field and Figure~\ref{fig:Figure4}(b) shows the spatial frequency of the reconstruct neural field from the basis function decomposition. The cut-off frequency of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing of the dominant dynamics.

@@

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(c). This band-limiting effect of the basis function decomposition can be observed when comparing Figure~\ref{fig:Figure4}(a) to Figure~\ref{fig:Figure4}(b).

@@

\subsubsection{Monte Carlo Simulation with Fixed Parameters}
A Monte Carlo approach was employed where 150 realisations of the data were generated. Each realisation consisted of 500~ms of data (sampled at $1$~kHz) and the estimation was applied to the final 400~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Ten iterations of the estimation algorithm were used to ensure the parameters converged. Figure~\ref{fig:Figure6} shows the rate of convergence, confirming the parameters converged, with the changes in the parameter errors falling to less than $10^{-4}$ after 6 iterations.

@@

% \tiny{\begin{tabular}{lcccccccccccc}
% Experiment&\multicolumn{4}{c}{I}&\multicolumn{4}{c}{II}&\multicolumn{4}{c}{III}\\
%  Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$&$\theta_0$&$\theta_1$&$\theta_2$&$\xi$\\
% \hline\hline
% Actual Value&$90$&$-80$&$5$&$0.9$&$100$&$-80$&$5$&$0.9$&$110$&$-80$&$5$&$0.9$\\
% Mean Estimate&$103.86$&$-89.05$&$5.23$&$0.923$& $101.75$&$-81.00$&$4.76$&$0.924$&$98.90 $&$-73.13$&$4.37$&$0.925$\\
% Std. Dev.&$23.14$&$16.57$&$0.74$&$0.003$ & $21.30$&$14.82$&$0.65$&$0.003$&$16.48$& $11.72$&$0.54$&$0.003$\\
% Bias (\%)& $15.4$&$11.31$&$4.6$ &$2.56$ &$1.75$&$1.25$&$4.8$&$2.67$&$10.1$&$ 8.58$&$12.6$&$2.78$  \\
% \end{tabular}\label{tab:MexicanHatKernelResults}}

\begin{table}[!ht]
	\caption{
	\bf{Results from varying the connectivity kernel parameters}}
\begin{tabular}{c|ccccc}
	Experiment & Parameter & Actual & Mean Est. & Std. Dev. & Bias (\%) \\ \hline
	I &	$\theta_0$ & $90$ & 103.86 & 23.14 & 15.4 \\
	& $\theta_1$ & -80 & -89.05 & 16.57 & -11.31 \\
	& $\theta_2$ & 5 & 5.23 & 0.74 & 4.6 \\
	& $\xi$ & 0.9 & 0.923 & 0.003 & 2.56 \\ \hline

	II & $\theta_0$ & $100$ & 101.75 & $21.30$ & $1.75$ \\
	& $\theta_1$ & -80 & -81.00 & $14.82$ & $1.25$ \\
	& $\theta_2$ & 5 & $4.76$ & $0.65$ & $4.8$ \\
	& $\xi$ & 0.9 & $0.924$ & $0.003$ & $2.67$ \\ \hline

	III & $\theta_0$ & $110$ & $98.90$ & $16.48$ & $10.1$ \\
	& $\theta_1$ & -80 & $-73.13$ & $11.72$ & $ 8.58$ \\
	& $\theta_2$ & 5 & $4.37$ & $0.54$ & $12.6$ \\
	& $\xi$ & 0.9 & $0.925$ & $0.003$ & $2.78$ \\ \hline
\end{tabular}\label{tab:MexicanHatKernelResults}
\end{table}
